\huge {\bold Parallel Algorithms}

### 2024-09-27 Papers/Parallel Aalgorithms/Booklet/PA.tex
Commit
Commit to end of Shared Memory SIMD Computers

@@

                    \item {\bold Shared memory:} All processors access a common memory space to fetch or store data. In Shared-Memory (SM) SIMD computers, the communication between processors occurs through a shared memory space. This model is often referred to as the Parallel Random-Access Machine (PRAM) model, where multiple processors work in parallel, and data exchange happens via common memory access.

@@

                    \begin{minipage}{0.49\linewidth} % First

                        \footnotesize

@@

                        \begin{tcolorbox}[colframe=white, size=small]

                            Phase 1:

@@

                            Phase 2:

@@

                            Initially, F is set to false. The first processor to find x sets F to true. All processors check F at each step, and if F becomes true, the search ends. The broadcast of F takes $O(\log N)$ time per step.
                            In the worst case, the EREW model requires $+ O(N\log N(N \log(\frac{N}{n})))$ time.

                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.49\linewidth} % Second

                        \footnotesize

@@

                    \end{minipage}

                    \begin{minipage}{0.48\linewidth} % Thired

                        \footnotesize

@@

                        \end{tcolorbox}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.49\linewidth} % Fourth

                        \footnotesize

                        \begin{tcolorbox}[colframe=white, size=small]

@@

                            Phase 1:

@@

@@

                    \end{minipage}

>>>>

            \subsection{SIMD Architecture}

@@

                    \item {\bold Shared memory:} All processors access a common memory space to fetch or store data.

                \subsubsection{{\bold > \hspace{10pt} Shared Memory (SM) SIMD Computers}}

                    In Shared-Memory (SM) SIMD computers, the communication between processors occurs through a shared memory space. This model is often referred to as the Parallel Random-Access Machine (PRAM) model, where multiple processors work in parallel, and data exchange happens via common memory access.

@@

                        \begin{minipage}{0.49\linewidth} % First
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

@@

                                {\bold Phase 1:}

@@

                                {\bold Phase 2:}

@@

                                Initially, F is set to false. The first processor to find x sets F to true. All processors check F at each step, and if F becomes true, the search ends. The broadcast of F takes $O(\log N)$ time per step.
                                In the worst case, the EREW model requires $+ O(N\log N(N \log(\frac{N}{n})))$ time.
                            \end{tcolorbox}
                        \end{minipage}
                        \hfill
                        \begin{minipage}{0.49\linewidth} % Second
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

@@

                            \end{tcolorbox}
                        \end{minipage}

                        \begin{minipage}{0.49\linewidth} % Thired
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

@@

                            \end{tcolorbox}
                        \end{minipage}
                        \hfill
                        \begin{minipage}{0.49\linewidth} % Fourth
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

@@

                                {\bold Phase 1:}

@@

                            \end{tcolorbox}
                        \end{minipage}

@@

                    When designing shared-memory SIMD computers, one of the main challenges is how to efficiently manage access to memory locations shared by multiple processors. The key concern is the cost of routing data between processors and memory, which scales with the number of processors $N$ and memory locations $M$. Now the problem overview raised is below:

                        \item Memory circuits need to be established to create pathways between the processors and the shared memory locations.

                        \item If each processor has its own dedicated memory, the cost for this approach is proportional to $f(M)$.

                        \item Sharing memory among processors reduces the total memory cost, but for large $N$ and $M$, the circuit cost can still become significant, growing proportionally to $N \times f(M)$.

                    To address this problem, several strategies have been proposed:

                        \item {\bold Approach 1:} Dividing shared memory into $R$ blocks of equal size, where each processor can access a block of memory at a time ($\frac{M}{R}$). This reduces the required memory per processor.

                        \item {\bold Approach 2:} Distributing $N$ memory locations equally among $N$ processors, with each processor handling a specific portion of the memory ($\frac{M}{N}$). This helps balance the workload and reduce memory contention.

                    \begin{tcolorbox}[colback=white, colframe=black!25, title= Example$:$ Memory Partitioning and Access (SM SIMD)]

                        In more detail, consider the scenario where memory is shared between $N$ processors by dividing the memory into $R$ blocks of equal size, where each processor can access one block at a time. For example, if $R = 3$ and $N = 5$, the processors share memory with each accessing its respective block.

                        \begin{minipage}{0.49\linewidth}
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

                                {\bold Key Points:}

                                    \item In each time step, a processor can only access one block of shared memory ($\frac{M}{R}$).

                                    \item The $N + R$ bidirectional links provide access between each processor and its designated block of memory.

                                    \item The overall circuit cost grows according to $N \times R \times f(\frac{M}{R})$, with an additional switch needed for managing simultaneous access.

                            \end{tcolorbox}
                        \end{minipage}
                        \hfill
                        \begin{minipage}{0.49\linewidth}
                            \footnotesize
                            \vspace{0.3cm}
                            \begin{tcolorbox}[colframe=white, size=small]

                                In practice, this model ensures efficient memory access without overwhelming the system with excessive memory demands, but it requires careful management of access timing and switching to avoid conflicts.

                            \end{tcolorbox}
                        \end{minipage}

                        \vspace{0.4cm}
                        \begin{center}
                            \begin{tikzpicture}

                                % Processors
                                \node[draw, minimum width=2.5cm, minimum height=1cm, line width=1] (P1) at (0,0) {\bold Processor 1};
                                \node[draw, minimum width=2.5cm, minimum height=1cm, line width=1] (P2) at (0,1.5) {\bold Processor 2};
                                \node[draw, minimum width=2.5cm, minimum height=1cm, line width=1] (P3) at (0,3) {\bold Processor 3};
                                \node[draw, minimum width=2.5cm, minimum height=1cm, line width=1] (P4) at (0,4.5) {\bold Processor 4};
                                \node[draw, minimum width=2.5cm, minimum height=1cm, line width=1] (P5) at (0,6) {\bold Processor 5};

                                % Memories
                                \node[draw, minimum width=3cm, minimum height=1cm, line width=1] (M1) at (3.5,-1.5) {\bold Memory Block 1};
                                \node[draw, minimum width=3cm, minimum height=1cm, line width=1] (M2) at (7,-1.5) {\bold Memory Block 2};
                                \node[draw, minimum width=3cm, minimum height=1cm, line width=1] (M3) at (10.5,-1.5) {\bold Memory Block 3};

                                % Joints
                                \foreach \i in {1,2,3,4,5} {
                                    \foreach \j in {1} {
                                        \draw[-] (P\i.east) -- ++(2.25,0) -- (M\j.north);
                                        }
                                        }

                                \foreach \i in {1,2,3,4,5} {
                                    \foreach \j in {2} {
                                        \draw[-] (P\i.east) -- ++(5.75,0) -- (M\j.north);
                                        }
                                        }

                                \foreach \i in {1,2,3,4,5} {
                                    \foreach \j in {3} {
                                        \draw[-] (P\i.east) -- ++(9.25,0) -- (M\j.north);
                                        }
                                        }

                                % Circles
                                \draw[black, fill=black] (3.5,0) circle (0.1cm);
                                \draw[black, fill=black] (7,0) circle (0.1cm);
                                \draw[black, fill=black] (10.5,0) circle (0.1cm);
                                \draw[black, fill=black] (3.5,1.5) circle (0.1cm);
                                \draw[black, fill=black] (7,1.5) circle (0.1cm);
                                \draw[black, fill=black] (10.5,1.5) circle (0.1cm);
                                \draw[black, fill=black] (3.5,3) circle (0.1cm);
                                \draw[black, fill=black] (7,3) circle (0.1cm);
                                \draw[black, fill=black] (10.5,3) circle (0.1cm);
                                \draw[black, fill=black] (3.5,4.5) circle (0.1cm);
                                \draw[black, fill=black] (7,4.5) circle (0.1cm);
                                \draw[black, fill=black] (10.5,4.5) circle (0.1cm);
                                \draw[black, fill=black] (3.5,6) circle (0.1cm);
                                \draw[black, fill=black] (7,6) circle (0.1cm);
                                \draw[black, fill=black] (10.5,6) circle (0.1cm);

                            \end{tikzpicture}        
                        \end{center}
                        \vspace{0.4cm}
